Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Sequence of events reviewed

The sequence of events is as follows:
\begin{enumerate}
\item First the {\sf NodIO} process is started in Node.js. As part of the
initialization procedure the sequential number of experiment is initialized and a log
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP
request and the page is loaded.
\item The main script renders the page and creates the worker instance. In
order to create a dedicated worker instance, another GET request is sent to the
server in order to receive the worker implementation script.
Each worker instance executes over time one or more evolutionary algorithms
{\em islands}. Each starts with a random population and it is assigned a
universally unique identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.
The initial parameters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following the previously
specified parameters. Once the population has been initialized, the EA
generation loop is started. After {\em n} generations the state of the population
is evaluated. First the chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not the global best,
the EA loop must continue, but first a chromosome is randomly selected from
the shared population. Also the current generation number and the best
chromosome is sent to the client's main thread in order to update the DOM to
display the current state of the island (described next). These messages are
all asynchronous, so even before a message is received by the target, the EA loop of
the worker continues. If the best fitness is a global best the best
chromosome is sent to the server and the loop ends. A message is sent to the
main thread indicating that the best individual has been found.
\item The main thread receives messages from workers as events handled by
a callback function. When an iteration message is received, the page is
updated by displaying a dynamic plot of the generation number and best
fitness, and a representation of the chromosome is also generated.
\item When a global best is received from an island the current experiment
ends, the experiment number is incremented and the population array is reset.
\item When the main thread receives a message indicating that a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the start-up cost, only the parameters and population are reset and a new UUID
assigned.
\end{enumerate}
%
\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%

>>>>

The sequence of events happens as follows:
\begin{enumerate}
\item First the {\sf NodIO} process is started in Node.js. As part of the
initialization procedure, the sequential number of experiments is initialized and a log
file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP
request, then the page is loaded.
\item The main script renders the page and creates the worker instance. In
order to create a dedicated worker instance, another GET request is sent to the
server, then the client receives the second worker implementation script.
Each worker instance executes over time one or more evolutionary algorithms
{\em islands}. Each island starts with a random population and it is assigned a
universally unique identifier (UUID). The UUID is included in each of the HTTP
requests sent to the server.
The initial parameters for each island
can be distinct. Once the worker is created and the parameters have been set,
a message is sent to the worker to start the algorithm.
\item The worker first creates the population following the previously
specified parameters. Once the population has been initialized, the EA
generation loop is started. After {\em n} generations the state of the population
is evaluated. The chromosome with the best fitness is sent to the shared
population with a PUT HTTP request. If the best fitness is not the global best,
the EA loop must continue, but first a chromosome is randomly selected from
the shared population. The current generation number and the best
chromosome are also sent to the client's main thread in order to update the DOM to
display the current state of the island (described next). These messages are
all asynchronous, so the EA loop of
the worker continues even before a message is received by the target. If the best fitness is a global best, the best
chromosome is sent to the server and the loop ends. A message is sent to the
main thread indicating that the best individual has been found.
\item The main thread receives messages from workers as events handled by
a callback function. When an iteration message is received, the page is
updated by displaying a dynamic plot of the generation number and best
fitness, and a representation of the chromosome is also generated.
\item When a global best is received from an island, the current experiment
ends, the experiment number is incremented, and the population array is reset.
\item When the main thread receives a message indicating that a worker has ended the EA
loop, that worker is reinitialized. The worker process is not ended, to
avoid the start-up cost, only the parameters and population are reset and a new UUID
assigned.
\end{enumerate}
%
\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
